---
title: "regen_revisits_analysis"
author: "CT"
date: "2023-06-29"
output: html_document
---

This script analyses post-fire revisited plot data from 7 fires in northern California to investigate trends in conifer seedling density and vegetation structure over time.

Analyses include: (a) modeling revisit density of conifer seedlings by genera, (b) visualizing establishment of seedlings by time since fire and year,  modeling change in density over time of (c) conifer seedlings and (d) vegetation components (grass, shrubs, forbs, and dominant shrub species)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(lme4)
library(MASS)
library(performance) #for checking overdispersion
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(patchwork)
library(viridis)
library(reshape2) 
library(reshape) 
library(GGally)
data_dir = readLines(here("data_dir.txt"), n=1)
source(here("scripts/convenience_functions.R"))
```

Load and clean revisited plot data
```{r}
load = read_csv(datadir("final_data/final_data_for_revisit_conifer_anlyses.csv"))

#combine hardwoods and shrub cover
df = load %>% mutate(Shrubs = coalesce(Shrubs, 0),
                     HARDWOOD = coalesce(HARDWOOD, 0),
                     Shrubs = Shrubs + HARDWOOD,
                     #replace NA and Inf distance to seed source with large number
                      DistPre = coalesce(DistPre, 400))

#replace DistPre inf with 400 
df$DistPre[is.infinite(df$DistPre)] <- 400

## add column for yes/no more seedlings at revisit and yes/no fewer seedlings at revisit
df = df %>% mutate(net_recruit_TF = factor(seedling_count_revisit > seedling_count_initial),
                   net_mort_TF = factor(seedling_count_revisit < seedling_count_initial))



#remove ANG - something strange with the plot codes -,  CLV - too long after fire and hardly any seedlings, and AMR - outlier fire with only one plot
df2 = df %>%
  filter(Fire != "Angora" & Fire != "CLV" & Fire != "American River Complex")
  
#clean up columns 
df_clean = df2 %>% 
  mutate(yrs2revisit = sample_yr_revisit - fire_year,
         seedling_count_initial_sqrt = sqrt(seedling_count_initial),
         stocked_initial = case_when(seedling_count_initial > 0 ~ 1,
                                     seedling_count_initial <= 0 ~ 0),
         stocked_revisit = case_when(seedling_count_revisit > 0 ~ 1,
                                     seedling_count_revisit <= 0 ~ 0),
         gained_stocking = case_when(stocked_initial == 0 & 
                                       stocked_revisit == 1 ~ 1,
                                     .default = 0),
         lost_stocking = case_when(stocked_initial == 1 & 
                                       stocked_revisit == 0 ~ 1,
                                     .default = 0),
         #change shrub cover so that it cannot exceed 100%
         Shrubs = case_when(Shrubs > 100 ~ 100, TRUE ~ Shrubs),
         #convert counts to density
         initial_density = seedling_count_initial/60,
         revisit_density = seedling_count_revisit/60,
         initial_density_scale_sqrt = sqrt(initial_density),
         #add vars for scaling in the next step
         yrs2initial_scale = yrs2initial,
         Shrubs_scale = Shrubs,
         time_between_samples_scale = time_between_samples,
         seedling_count_initial_scale = seedling_count_initial_sqrt,
         DistPre_scale = DistPre,
         heat_load_scale = heat_load,
         )

#write_csv(df_clean, datadir("revisit_analysis_cleaned_data/cleaned_data_from_analysis.csv"))

```


Explore data summaries
```{r}
#summarise data
sum_table = df_clean %>% group_by(Fire, sp_func) %>%
  summarise(seedling_count_initial = length(Plot_ID[seedling_count_initial > 0]),
            seed_source50 = length(Plot_ID[DistPre <= 50]))

sum_table2 = df_clean %>% group_by(Plot_ID) %>%
  summarise(seedling_count_initial = sum(seedling_count_initial),
            seedling_count_revisit = sum(seedling_count_revisit))


plots_per_fire = df_clean %>% filter(sp_func == "pine") %>%
  group_by(Fire) %>%
  summarise(n = n())

#79 plots
fire_table = df_clean %>% dplyr::select(Fire, fire_year,  initial_sample_yr, yrs2initial, sample_yr_revisit, time_between_samples) %>%
  distinct() %>%
  mutate(total_time = sample_yr_revisit - fire_year)


#summarise elev data
elev_sum = df_clean %>% filter(sp_func == "pine") %>%
  group_by(Fire) %>%
  summarise(max = max(elev),
            min = min(elev),
            mean = mean(elev))

```


Explore data correlations

>tmean and temp highly correlated ~.5
>very zero inflated? Check for overdispersion after fitting the model

```{r}
ggplot(df_clean, aes(seedling_count_revisit, fill = sp_func)) + geom_histogram(binwidth = 1) + facet_grid(sp_func ~ 
    ., margins = TRUE, scales = "free")

ggplot(df_clean %>% filter(sp_func == "pine"), aes(yrs2initial, fill = Fire)) + geom_histogram(binwidth = 1)+theme_bw(13)

ggplot(df_clean %>% filter(sp_func == "pine"), aes(time_between_samples, fill = Fire)) + geom_histogram(binwidth = 1)+theme_bw(13)


```



```{r}
ggpairs(df_scaled[, c("seedling_count_initial", "seedling_count_revisit", "DistPre", "yrs2initial", "time_between_samples", "Shrubs", "precip", "tmean", "twi", "heat_load", 'seedling_count_initial_sqrt')], progress = FALSE)
```

> Precip and temp are highly correlated


#scale data for revisit analysis
```{r}
df_scaled = df_clean %>% mutate_at(c('DistPre_scale', 'Shrubs_scale', 'heat_load_scale','seedling_count_initial_scale', 'yrs2initial_scale', 'time_between_samples_scale'), ~(scale(.) %>% as.vector))

#separate SIM from other Sierra fires for modeling
df_scaled_nosim = df_scaled %>% filter(Fire != "SIM")
df_scaled_sim = df_scaled %>% filter(Fire == "SIM") #separate SIM because it's the only fire with substantial PSME and not much else.

#create df with all species lumped
lumped_df = df_clean %>% group_by(Plot_ID, Fire, heat_load, Shrubs,  time_between_samples, yrs2initial) %>% 
  summarise(DistPre = min(DistPre),
            seedling_count_initial = sum(seedling_count_initial, na.rm = TRUE),
            seedling_count_initial_sqrt = sqrt(sum(seedling_count_initial, na.rm = TRUE)),
            seedling_count_revisit = sum(seedling_count_revisit, na.rm = TRUE),
            net_seedling_dif = seedling_count_revisit - seedling_count_initial)


#scale lumped df
lumped_df_scale = lumped_df %>% data.frame() %>% mutate_at(c('DistPre', 'Shrubs', 'heat_load',  'seedling_count_initial_sqrt', 'yrs2initial', 'time_between_samples'), ~(scale(.) %>% as.vector))
```

# Explore covariate relationships with seedling density
```{r}

tmp <- reshape2::melt(df_clean[, c("net_recruit_TF", "seedling_count_initial", "DistPre", "time_between_samples", "Shrubs")],
  id ="net_recruit_TF")

tmp <- reshape2::melt(df_clean[, c("net_recruit_TF", "seedling_count_initial", "DistPre", "time_between_samples", "Shrubs")],
  id ="net_recruit_TF")

tmp2 <- reshape2::melt(df_clean[, c("net_mort_TF", "seedling_count_initial", "DistPre", "time_between_samples", "Shrubs")],
  id ="net_mort_TF")



ggplot(tmp, aes(net_recruit_TF, y = value, fill=net_recruit_TF)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free_y")

ggplot(tmp2, aes(net_mort_TF, y = value, fill=net_mort_TF)) +
  geom_boxplot() +
  facet_wrap(~variable, scales="free_y")


```

Plot relationsips between initial and revisit seedlings with 1:1 lines
```{r} 
pal =  c(
  "#00798c",
  "#d1495b",
  "#edae49",
  "#66a182",
  "#2e4057",
  "#8d96a3",
  "#6F4685"
)



(p_all = ggplot(df_scaled %>% filter(sp_func == "fir" | sp_func == "Douglas-fir" | sp_func == "pine") %>%
                  mutate(Fire = case_when(Fire == "GON" ~ "Gondola", Fire == "MNL" ~ "Moonlight", Fire == "SHR" ~ "Showers", Fire == "SIM" ~ "Sims", TRUE ~ Fire)), 
  aes(x = sqrt(seedling_count_initial/60), y = sqrt(seedling_count_revisit/60), color = Fire))+
  geom_point(size = 2.8, alpha = 0.8) +
  theme_bw(13)+
  geom_abline(slope = 1)+
        scale_color_manual(values = pal)+
    # geom_smooth(method = "lm")+
  xlab(bquote(`initial seedlings m`^`-2`))+
  ylab(bquote(`re-survey seedlings m`^`-2`))+
  scale_y_continuous(labels = function(x) x^2, breaks = c(0, 0.5, sqrt(.75), sqrt(1.5)))+
  scale_x_continuous(labels = function(x) x^2, breaks = c(0, 0.5, sqrt(.75), sqrt(1.5)))+
  facet_grid(~sp_func))



# ggsave(datadir("figures/raw_counts_cse_all_fires.png"), width = 9, height = 3.5)
```



# Data Analysis


## (a) Model revisit seedling density as a function of initial seedling density and covariates


Pine revisits model
```{r}

m_pine = glmmTMB(seedling_count_revisit ~ seedling_count_initial_scale + Shrubs_scale + heat_load_scale + DistPre_scale + yrs2initial_scale + time_between_samples_scale, data = df_scaled_nosim %>% filter(sp_func == "pine"), family = nbinom2)


#check model assumptions
check_overdispersion(m_pine)
simoutput = simulateResiduals(m_pine, re.form = NULL)
testDispersion(simoutput)

#create pine only df for plotting
df_pine = df_scaled_nosim %>% filter(sp_func == "pine")
plotResiduals(m_pine, form = df_pine$seedling_count_initial_scale)

summary(m_pine)


#modele seedling count revisit without seedling count initial to compute partial r2
m_pine_ns = glmmTMB(seedling_count_revisit ~  Shrubs_scale + heat_load_scale + DistPre_scale + yrs2initial_scale + time_between_samples_scale, data = df_scaled_nosim %>% filter(sp_func == "pine"), family = nbinom2)

```

Fir revisits model
```{r}
m_fir = glmmTMB(seedling_count_revisit ~ seedling_count_initial_scale + Shrubs_scale + heat_load_scale + DistPre_scale + yrs2initial_scale + time_between_samples_scale, data = df_scaled_nosim %>% filter(sp_func == "fir"), family = nbinom2)

#test assumptions
simoutput = simulateResiduals(m_fir, re.form = NULL)
testDispersion(simoutput)
df_fir = df_scaled_nosim %>% filter(sp_func == "fir")
plotResiduals(m_fir, form = df_fir$seedling_count_initial_scale)
plotResiduals(m_fir)

summary(m_fir)

#model without initial seedling count to compute r2
m_fir_ns = glmmTMB(seedling_count_revisit ~ Shrubs_scale + heat_load_scale + DistPre_scale + yrs2initial_scale + time_between_samples_scale, data = df_scaled_nosim %>% filter(sp_func == "fir"), family = nbinom2)
```

Douglas-fir revisits model - only data from SIMs fire
```{r}
m_sim = glmmTMB(seedling_count_revisit ~ seedling_count_initial_scale  + Shrubs_scale  + heat_load_scale  + DistPre_scale , data = df_scaled_sim %>% filter(sp_func == "Douglas-fir"), family = nbinom2)

#test assumptions
simoutput = simulateResiduals(m_sim, re.form = NULL)
testDispersion(simoutput)
df_fir_sim = df_scaled_sim %>% filter(sp_func == "Douglas-fir")
plotResiduals(m_sim, form = df_fir_sim$seedling_count_initial_scale)
plotResiduals(m_sim) 
summary(m_sim)

# without initial seedlings
m_sim_ns = glmmTMB(seedling_count_revisit ~ Shrubs_scale  + heat_load_scale  + DistPre_scale, data = df_scaled_sim %>% filter(sp_func == "Douglas-fir"), family = nbinom2)

#compute r2
r2_nakagawa(m_sim)
r2_nakagawa(m_fir)
r2_nakagawa(m_pine)
```

make plots for revisit density models 
```{r}
pal = c("#07393C", "#2C666E", "#A4AC96")
pal2 = c("#2C666E", "#A4AC96")

(p1 = plot_models(m_pine, m_fir, m_sim,
            axis.labels = rev(c('initial seedling count',
                            'shrub cover', 
                            'heat load',
                            'dist. to seed source',
                            'time to initial sample',
                            'time between samples'
                            )),
            m.labels = c("pine", "fir", "Douglas-fir"),
            axis.lim = c(0.1,10),
            legend.title = "",
            colors = pal) + 
  theme_bw(14)+
  geom_hline(yintercept = 1, linetype = 2)
)

#ggsave(datadir('figures/revisit_count_model_all_fires.png'), height = 4, width = 5)

```



### Compute partial r2 for seedling models
```{r}
#compute partial R2 for initial seedling count (sqrt)
r2_nakagawa(m_fir)$R2_marginal - r2_nakagawa(m_fir_ns)$R2_marginal #0.00133
r2_nakagawa(m_pine)$R2_marginal - r2_nakagawa(m_pine_ns)$R2_marginal # 0.3052775 
r2(m_sim)$R2_marginal - r2(m_sim_ns)$R2_marginal #0.4602
```


Model probability of a plot having net mortality & net recruitment of all conifers between the two visits
```{r}
#create binary vairable for net mortality and net recruitment
df_mort = lumped_df %>% mutate(net_mortality = case_when(net_seedling_dif < 0 ~ 1, TRUE ~ 0),
                               net_recruitment = case_when(net_seedling_dif > 0 ~ 1, TRUE ~ 0))

#model drivers of net mortality and recruitment
m_mort = glmmTMB(net_mortality ~ seedling_count_initial +  Shrubs + heat_load + DistPre + yrs2initial + time_between_samples + (1|Fire), data = df_mort,
                 family = "binomial")



m_recr = glmmTMB(net_recruitment ~ Shrubs + heat_load + DistPre + yrs2initial + time_between_samples + (1|Fire), data = df_mort,
                 family = "binomial")

#plot residuals and summarise model output
plotResiduals(m_mort) 
summary(m_mort)
plot_model(m_mort)

plotResiduals(m_recr) 
summary(m_recr)

#compare means with emmeans
emmeans(m_mort, ~ Shrubs, at = list(Shrubs = 10), type = "response")
emmeans(m_mort, ~ seedling_count_initial, at = list(seedling_count_initial = 10), type = "response")


#plot models
(pa = plot_model(m_mort, type = "pred", terms = "Shrubs [all]", colors = "gs") +
    theme_bw(15)+
    theme( panel.grid.minor = element_blank())+
    ggtitle("Pr(net mortality)")+
    xlab("perc. shrub cover")+
    geom_rug(alpha = .5, sides = "b")
  )


(pb = plot_model(m_recr, type = "pred", terms = "heat_load [all]", colors = "gs") + theme_bw(15) +
    theme( panel.grid.minor = element_blank())+
    ggtitle("Pr(net recruitment)")+
    geom_rug(alpha = .5, sides = "b")
  )


(pc = plot_model(m_recr, type = "pred", terms = "DistPre [all]", colors = "gs") + theme_bw(15) +
    xlab("dist. to seed source (m)")+
    theme( panel.grid.minor = element_blank())+
    ggtitle("Pr(net recruitment)")+
    geom_rug(alpha = .5, sides = "b")
  )


#combine plots
pa + pb + pc + plot_annotation(tag_levels = 'a')

# ggsave(datadir('figures/recruit_mortality_marginal_effects.png'), height = 4, width = 12)

```




## (b) Explore establishment patterns from initila and revisit- seedling ages are not fully reliable given difficulty aging older seedlings
```{r}
#read in initial seedling age data
initial_seed_age = read_csv(datadir("final_data/initial_budscar_compiled.csv")) %>%
  mutate(Year = sample_yr)

#remove plots that are not included in the cleaned revisit data and the revisit models above
initial_seed_age_sub = initial_seed_age %>% filter(Plot_ID %in% df_clean$Plot_ID)

#check that all plots are accounted for
# t = df_clean %>% filter(!Plot_ID %in% initial_seed_age$Plot_ID)
#all accounted for except plots with no initial seedlings

#read in revisit seedling age data
revisits_seed_age = read_csv(datadir("final_data/revisits_budscar_compiled.csv"))

#remove plots that are not included in the cleaned revisit data and the revisit models above
revisits_seed_age = revisits_seed_age %>% filter(Plot_ID %in% df_clean$Plot_ID)

#check that all plots are accounted for
# t = df_clean %>% filter(!Plot_ID %in% revisits_seed_age$Plot_ID)
#all accounted for except plots with no revisit seedlings

seed_age = bind_rows(revisits_seed_age, initial_seed_age_sub) %>%
  dplyr::select(Plot_ID, Species, BudScars, Year, Visit) %>%
  mutate(Plot_ID = toupper(Plot_ID)) %>% 
  mutate(sp_func = Species,
         sp_func = sub("ABCO/PSME", "fir", sp_func), #adding psme to firs in Sierra plots
         sp_func = sub("ABIES", "fir", sp_func), #combining all firs
         sp_func = sub("ABCO", "fir", sp_func), 
         sp_func = sub("ABMA", "fir", sp_func), 
         sp_func = sub("CADE", "cedar", sp_func),
         sp_func = sub("PICO", "pine", sp_func), #combine all pines
         sp_func = sub("PIMO", "pine", sp_func),
         sp_func = sub("PIPJ", "pine", sp_func),
         sp_func = sub("PINUS", "pine", sp_func),
         sp_func = sub("PILA", "pine", sp_func),
         sp_func = sub("PSME", "Douglas-fir", sp_func),
         sp_func = sub("ARME", "hrdwd", sp_func), #lump hardwoods
         sp_func = sub("LIDE", "hrdwd", sp_func),
         sp_func = sub("QUCH", "hrdwd", sp_func),
         sp_func = sub("QUKE", "hrdwd", sp_func),
         sp_func = sub("SASC", "hrdwd", sp_func),
         sp_func = sub("CONU", "hrdwd", sp_func),
         sp_func = sub("QUGA", "hrdwd", sp_func),
         sp_func = sub("ALRH", "hrdwd", sp_func),
         sp_func = sub("COCO", "hrdwd", sp_func),
         sp_func = sub("SALA", "hrdwd", sp_func),
         sp_func = sub("SALIX", "hrdwd", sp_func),
         sp_func = sub("SANI", "hrdwd", sp_func),
         sp_func = sub("SANIC", "hrdwd", sp_func)) %>%
  filter(sp_func == "fir" | sp_func == "pine" | sp_func == "Douglas-fir")



#clean seed establishment data
seed_age_sub = seed_age %>% filter(Plot_ID %in% df_clean$Plot_ID) %>%
  left_join(df_clean %>% filter(sp_func == "pine") %>% dplyr::select(Plot_ID, Fire, fire_year)) %>%
  mutate(yrs2fire = Year - fire_year,
         post_fire_estab = yrs2fire - BudScars,
         #for one sim plot where establishment year is -1, change to 0
         BudScars = case_when(post_fire_estab < 0 ~ BudScars - 1,
                             TRUE ~ BudScars),
         post_fire_estab = yrs2fire - BudScars,
         estab_year = Year - BudScars)

#change DF to fir in all plots except Sims because Douglas-fir was not recorded in any initial sample
seed_age_sub = seed_age_sub %>%
  #change DF to fir in all plots except Sims
  filter(Fire != "SIM") %>%
  mutate(sp_func = sub("Douglas-fir", "fir", sp_func)) %>%
  bind_rows(seed_age_sub %>% filter(Fire == "SIM"))

# #test to make sure counts match cleaned df
# t2 = df_clean %>% group_by(Fire, sp_func) %>%
#   summarise(initial = sum(seedling_count_initial),
#             revisit = sum(seedling_count_revisit))

```


Scale by number of sample plots - time since fire establishment
```{r}
#get plots per fire for scaling
plots_per_fire = lumped_df %>%
  group_by(Fire) %>%
  summarise(plots_per_fire = n())


#create dfs with plot counts by establishment year and time since fire
seed_age_y2f_scale = seed_age_sub  %>%
  group_by(Fire, Visit, post_fire_estab, sp_func) %>%
  summarise(count_regen = n()) %>%
  left_join(plots_per_fire) %>%
  mutate(count_per_plot = count_regen/plots_per_fire)

seed_age_y2f_scale_sum = seed_age_y2f_scale %>% group_by(Visit, post_fire_estab, sp_func) %>%
  summarise(count_regen_scale = sum(count_per_plot))

#plot separately for each fire
(pa = ggplot(seed_age_y2f_scale %>%
               mutate(Fire = factor(Fire, levels = c('SIM', 'SHR', 'Power',
                                                     'Pendola', 'MNL', 'GON', 'Freds')),
                      sp_func = factor(sp_func, levels = c('pine', 'fir', 'Douglas-fir')),
                      Visit = case_when(Visit == "revisit" ~ "resurvey",
                                        TRUE ~ Visit)), 
             aes(y = sqrt(count_per_plot), x = post_fire_estab,  color = Visit)) + 
    # geom_bar(position = "dodge") + 
     geom_line(linewidth = .7)+
    geom_point(size = 2.0)+
    theme_bw(13) +
  scale_color_manual(values = c("grey", "#2d4654"))+
    scale_y_continuous(labels = function(x) x^2)+
  labs(color='sample') +
  facet_grid(Fire~sp_func)+
  xlab("no. of years after fire seedlings established")+
  ylab("seedlings per plot")+
  theme(strip.background = element_rect(fill = "white")))

# ggsave(datadir('figures/estab_plot_separate_fires.svg'), width = 9, height = 6)

```


```{r}

#create dfs with plot counts by establishment year
seed_age_estabyr_scale = seed_age_sub %>% 
  group_by(Fire, Visit, estab_year, sp_func) %>%
  summarise(count_regen = n()) %>%
  left_join(plots_per_fire) %>%
  mutate(count_per_plot = count_regen/plots_per_fire)

seed_age_estabyr_scale_sum = seed_age_estabyr_scale %>% group_by(Visit, estab_year, sp_func) %>%
  summarise(count_regen_scale = sum(count_per_plot))

### PLOT
#establishment year by fire
(pb = ggplot(seed_age_estabyr_scale %>%
               mutate(Visit = case_when(Visit == "revisit" ~ "re-sample",
                                        TRUE ~ Visit)), aes(y = sqrt(count_per_plot), x = estab_year,  color = Visit)) + 
    # geom_bar(position = "dodge") + 
     geom_line(linewidth = .7)+
    geom_point(size = 2.0)+
    theme_bw(13) +
  scale_color_manual(values = c("grey", "#2d4654"))+
  labs(color='sample')+
  scale_y_continuous(labels = function(x) x^2)+
  facet_grid(Fire~sp_func)+
  xlab("post-fire establishment year")+
  ylab("seedlings per plot")+
    annotate(geom = "rect", 
           xmin = 2012,
                xmax = 2016,
                ymin = -Inf,
                ymax = Inf,
            alpha = 0.2, fill = "red")+    
  theme(strip.background = element_rect(fill = "white")))

# ggsave(datadir('figures/drought_plot_separate_fires.svg'), width = 9, height = 6)

```


## (c) Examine change in conifer density over time by genus
```{r}
df_density = df_clean %>% mutate(sp_func = case_when(Fire == "SIM" & sp_func == "fir" ~ "Douglas-fir", TRUE ~ sp_func )) %>%
  #change pendola fire ids to match each other - these are the same fire
  mutate(FIRE_ID = case_when(FIRE_ID == 'PNT' | FIRE_ID == 'PNP' ~ 'PEN', TRUE ~ FIRE_ID)) %>%
  dplyr::select(FIRE_ID, Plot_ID, initial_density, revisit_density, sp_func) %>%
  #convert to long format for boxplots
  pivot_longer(initial_density:revisit_density, names_to = 'visit', values_to = 'density')

#plot
ggplot(df_density %>%
         mutate(ecoregion = case_when(FIRE_ID == "SIM" ~ "Klamath Mnts.",
                                      TRUE ~ "Sierra Nevada")),
       aes(x = sp_func, y = sqrt(density), fill = visit))+
  geom_boxplot()+
  # stat_summary(fun.y = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y..),
  #              width = 0.75, size = 1, linetype = "solid")
  scale_fill_manual(values = c("grey", "#425865"))+
  theme_bw(13)+
  scale_y_continuous(labels = function(x) x^2)+
  xlab("")+
  ylab(bquote(`seedlings m`^`-2`))+
  facet_wrap(~ecoregion)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#summarise
summary = df_density %>% group_by(sp_func, visit) %>%
  summarise(mean = mean(density))
  
#create a summary of plot level conifer seedling density  
df_density_plot_level = df_density %>%
  group_by(Plot_ID, FIRE_ID, visit) %>%
  summarise(density_total = sum(density)) %>%
  mutate(seedling_presence = density_total > 0)

table(df_density_plot_level$seedling_presence, df_density_plot_level$visit)

#join to plot predictor data
df_density = df_density %>% left_join(df_clean %>% dplyr::select(Plot_ID, Fire, time_between_samples, yrs2initial) %>% distinct())

#pivot wider for computing differences
df_density_wide = df_density %>% pivot_wider(names_from = visit, values_from = density) %>%
  mutate(density_diff = revisit_density - initial_density)

df_density_diff = df_density%>% filter(Fire != "Sim") %>% group_by(visit, sp_func) %>%
  summarise(dif = sum(density))


```

#### Model changes in conifer density between visits
```{r}
#model change in pine density between visits
m_pden= glmmTMB(density ~ visit + time_between_samples + yrs2initial + (1|Plot_ID), data = df_density %>% filter(sp_func == "pine") %>% filter(Fire != "SIM"), family = tweedie)

plotResiduals(m_pden) 
summary(m_pden)

#model change in fir density between visits
m_fden= glmmTMB(density ~ visit + time_between_samples + yrs2initial + (1|Plot_ID), data = df_density %>% filter(sp_func == "fir") %>% filter(Fire != "SIM"), family = tweedie)

plotResiduals(m_fden) 
summary(m_fden)

#model change in douglas fir density between visits
m_dfden= glmmTMB(density ~ visit + (1|Plot_ID), data = df_density %>% filter(sp_func == "Douglas-fir") %>% filter(Fire == "SIM"), family = tweedie)

plotResiduals(m_dfden) 
summary(m_dfden)
```

compare conifer density means with emmeans
```{r}
#emmeans table
em_dfden = emmeans(m_dfden, specs = revpairwise ~ visit, data = df_density %>% filter(sp_func == "Douglas-fir") %>% filter(Fire == "SIM"), type = "response")

em_fden = emmeans(m_fden, specs = revpairwise ~ visit, data = df_density %>% filter(sp_func == "fir") %>% filter(Fire != "SIM"), type = "response")

#at = list(yrs2initial = c(6,10))

em_pden = emmeans(m_pden, specs = revpairwise ~ visit, data = df_density %>% filter(sp_func == "pine") %>% filter(Fire != "SIM"), type = "response")


#combine emmeans into table
bind_rows(data.frame(em_pden$contrasts), data.frame(em_fden$contrasts), 
          data.frame(em_dfden$contrasts))
```



## (d) Explore trends in shrubs and veg structure!
```{r}
#load in shrub species cover data
shrub_df = read_csv(datadir("final_data/shrub_data_compiled.csv")) %>%
  #change PNT and PNP to 'Pendola' these are from the same fire
  mutate(Fire = case_when(Fire == "PNT" | Fire == "PNP" ~ "Pendola", 
                          Fire == "PWR" ~ "Power",
                          Fire == "SIMS" ~ "SIM", 
                          .default = Fire))

#keep plots included in revisit conifer analysis
# i = df_clean$Plot_ID %in% shrub_df$PlotID 
# df_clean$Plot_ID[!i] #missing 2 freds plots: gon and sim plot have no shrubs in initial visit

# i = shrub_df$PlotID %in% df_clean$Plot_ID 
# shrub_df$PlotID[!i]

#keep only plots in the cleaned data with regen data
shrub_df_sub = shrub_df %>% filter(PlotID %in% df_clean$Plot_ID)

#make sure each plot is listed twice
t = shrub_df_sub %>% dplyr::select(PlotID, Sample_year, Fire) %>%
  distinct() %>%
  group_by(PlotID) %>%
  summarise(n = n())
#Sim18 and Gon06 had no shrub cover in initial visit


shrub_sp = shrub_df %>% dplyr::select(Species) %>% distinct()

#remove occurrences with less than 10% cover - some revisited plots only count shrub cover if >10%
shrub_dom = shrub_df_sub %>% 
  filter(Cover >=10) %>%
  #lump species by genus
  mutate(genus = substr(Species, 1,2))

shrub_fun = shrub_dom %>% group_by(Fire, visit, PlotID, Sample_year, genus) %>%
  summarise(Cover = sum(Cover))

#convert to wide and then back to get differences
shrubw = shrub_fun %>% pivot_wider(id_cols = -c(Sample_year, -Fire), names_from = visit, values_from = Cover) %>%
  #replace nas with 0s
  mutate(initial = ifelse(is.na(initial), 0, initial),
         revisit = ifelse(is.na(revisit), 0, revisit),) %>%
  mutate(dif = revisit - initial,
         Plot_ID = PlotID)

#add sample years
shrub_yr = left_join(shrubw, df_clean %>% dplyr::select(Plot_ID, initial_sample_yr, sample_yr_revisit, Fire, yrs2initial) %>% distinct()) %>%
  mutate(time_between_samples = as.numeric(sample_yr_revisit) - as.numeric(initial_sample_yr)) %>%
  #rename genus for ploting
  mutate(genus = case_when(genus == "AR" ~ "Arctostaphylos",
                           genus == "CE" ~ "Ceanothus",
                           genus == "CH"~ "Chamaebatia",
                           genus == "TO" ~ "Toxicodendron",
                           TRUE ~ "Other"),
         ecoregion = case_when(Fire == "SIM" ~ "Klamath Mnts.",
                               TRUE ~ "Sierra Nevada")) %>%
  #remove other species
  filter(genus != "Other")

#PLOT
(shrub_p = ggplot(shrub_yr, aes(x = genus, y = dif/time_between_samples))+
    geom_boxplot()+
    geom_jitter(alpha = 0.6,size = 3, aes(color = Fire), width = .25)+
  theme_bw(13)+
  ylab("annual change in shrub cover")+
  xlab(""))
  #scale_color_viridis(discrete = TRUE, option = "D"))
  # scale_color_manual(values = pal)

(p_shrub = ggplot(shrub_yr %>% 
                    pivot_longer( cols = revisit:initial, names_to = 'visit', values_to = 'cover') %>%
                    filter(genus != "Chamaebatia") %>%
                    mutate(visit = case_when(visit == "revisit" ~ "resurvey", TRUE ~ visit)), 
                  aes(x = genus, y = cover, fill = visit))+
    geom_boxplot()+
  theme_bw(13)+
      scale_fill_manual(values = c("grey", "#425865"))+
  xlab("")+
  facet_wrap(~ecoregion)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)))

p_shrub + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  #scale_color_viridis(discrete = TRUE, option = "D"))
  # scale_color_manual(values = pal)

#ggsave(datadir("figures/shrub_revisit_cse.png"), width = 6, height = 4, units = "in")
```

### Model differences in shrub species cover
```{r}
#convert to long format for subsetting by species
shrub_l = shrub_yr %>% dplyr::select(-dif) %>%
  pivot_longer(initial:revisit, names_to = 'visit', values_to = 'cover')

## Model arctostaphylos change over time
m_arc = glmmTMB(cover ~ visit  + time_between_samples + yrs2initial + (1|Plot_ID), data = shrub_l %>% filter(genus == "Arctostaphylos") %>% filter(Fire != "SIM"), family = tweedie)

#check model assumptions
plotResiduals(m_arc) 
summary(m_arc)

## Model ceanothus change over time
m_ce = glmmTMB(cover ~ visit  + time_between_samples + yrs2initial + (1|Plot_ID), data = shrub_l %>% filter(genus == "Ceanothus") %>% filter(Fire != "SIM"), family = tweedie)

plotResiduals(m_ce) 
summary(m_ce)

#not enough data points to model chamaebatia
# m_ch = glmmTMB(cover ~ visit  + time_between_samples + yrs2initial + (1|Plot_ID), data = shrub_l %>% filter(genus == "Chamaebatia") %>% filter(Fire != "SIM"), family = tweedie)
# 
# plotResiduals(m_ch) 
# summary(m_ch)


#Repeat models for Klamath ecoregion
## Model poison oak change over time
m_tox = glmmTMB(cover ~ visit + (1|Plot_ID), data = shrub_l %>% filter(genus == "Toxicodendron") %>% filter(Fire == "SIM"), family = tweedie)

plotResiduals(m_tox) 
summary(m_tox)


#model ceanothus for Klamath
m_ce_sim = glmmTMB(cover ~ visit + (1|Plot_ID), data = shrub_l %>% filter(genus == "Ceanothus") %>% filter(Fire == "SIM"), family = tweedie)

plotResiduals(m_ce_sim) 
summary(m_ce_sim)

```

Compare model means with emmeans 
```{r}
em_arc = emmeans(m_arc, specs = revpairwise ~ visit , data = shrub_l %>% filter(genus == "Arctostaphylos") %>% filter(Fire != "SIM"), type = "response")

em_ce = emmeans(m_ce, specs = revpairwise ~ visit , data = shrub_l %>% filter(genus == "Ceanothus") %>% filter(Fire != "SIM"), type = "response")

em_tox = emmeans(m_tox, specs = revpairwise ~ visit , data = shrub_l %>% filter(genus == "Toxicodendron") %>% filter(Fire == "SIM"), type = "response")

em_ce_sim = emmeans(m_ce_sim, specs = revpairwise ~ visit , data = shrub_l %>% filter(genus == "Ceanothus") %>% filter(Fire == "SIM"), type = "response")

#combine emmeans into table
bind_rows(data.frame(em_arc$contrasts), data.frame(em_ce$contrasts), 
          data.frame(em_tox$contrasts), data.frame(em_ce_sim$contrasts))
```



### Examine change in vegetation structrue over time
```{r}
#read in veg structure data
lf = read_csv(datadir("final_data/lifeform_compiled.csv"))

#remove gondola and showers because they don't have adequate initial visit data
lf_sub = lf %>% filter(Fire != "GON" & Fire != "SHR") %>%
  #keep only plots included in revisits conifer analyses
  filter(Plot_ID %in% df_clean$Plot_ID)

#make sure there are 2 records of each plot (initial and revisit)
t = lf_sub %>% dplyr::select(Plot_ID, Fire, visit) %>%
  distinct() %>%
  group_by(Plot_ID) %>%
  summarise(n = n())

#covert to long format so that all lifeforms are in one column
lf_long = lf_sub %>% 
  pivot_longer(Grams:Forbs, names_to = "lifeform", values_to = "cover") %>%
    mutate(cover = ifelse(is.na(cover), 0, cover))

#pivot wider so that cover at each visit has its own column
lf_wide = lf_long %>% dplyr::select(-Sample_year, -Fire) %>%
  pivot_wider(names_from = 'visit', values_from = "cover") %>%
  #remove NAs
  mutate(initial = ifelse(is.na(initial), 0, initial),
         revisit = ifelse(is.na(revisit), 0, revisit)) %>%
  #add back in Fire and sample years
  left_join(df_clean %>% dplyr::select(Plot_ID, initial_sample_yr, sample_yr_revisit, Fire, yrs2initial, time_between_samples) %>% distinct())

#join to cleaned data for plot level covariates
lf_long_join = left_join(lf_long, df_clean %>% dplyr::select(Plot_ID, initial_sample_yr, sample_yr_revisit, Fire, yrs2initial, time_between_samples) %>% distinct()) %>%
  mutate(ecoregion = case_when(Fire == "SIM" ~ "Klamath Mnts.",
                               TRUE ~ "Sierra Nevada"))

#summarise
# lf_long %>% filter(Fire == "SIM", 
#                    lifeform == "Shrubs",
#                    visit == "revisit") %>%
#   summary()
```

#### Plot structural compotents
```{r}
(p_lifeform = ggplot(lf_long_join %>%
                       mutate(visit = case_when(visit == "revisit" ~ "resurvey", TRUE ~ visit))
                     , aes(x = lifeform, y = sqrt(cover), fill = visit))+
  geom_boxplot()+
    scale_fill_manual(values = c("grey", "#425865"))+
  theme_bw(13)+
  xlab("")+
  facet_wrap(~ecoregion))

# ggsave(datadir("figures/lifeform_all.png"), width = 6, height = 4, units = "in")

(p_lifeform +
  theme(legend.title=element_blank())) / (p_shrub +
  theme(legend.title=element_blank())) +plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect")

# ggsave(datadir("figures/lifeform_and_shrub.png"), width = 6, height = 7, units = "in")
```

### model difference in veg structural component cover by ecoregion
```{R}
#convert cover to proportion for modeling
lf_long_join = lf_long_join %>%
  mutate(cover_prop = cover/100)

## Model lifeform change over time for Klamath/Sims fire
#model shrubs change over time
m_shrub_sim = glmmTMB(cover ~ visit  + (1|Plot_ID), data = lf_long_join %>% filter(lifeform == "Shrubs") %>% filter(ecoregion == "Klamath Mnts."), family = tweedie)

plotResiduals(m_shrub_sim) 
summary(m_shrub_sim)

#model grams change over time
m_gram_sim = glmmTMB(cover ~ visit  + (1|Plot_ID), data = lf_long_join %>% filter(lifeform == "Grams") %>% filter(ecoregion == "Klamath Mnts."), family = tweedie)

plotResiduals(m_gram_sim) 
summary(m_gram_sim)

#model forbs change over time
m_forb_sim = glmmTMB(cover ~ visit  + (1|Plot_ID), data = lf_long_join %>% filter(lifeform == "Forbs") %>% filter(ecoregion == "Klamath Mnts."), family = tweedie)

plotResiduals(m_forb_sim) 
summary(m_forb_sim)

## Repeat for Sierra fires
#model shrubs change over time
m_shrub = glmmTMB(cover ~ factor(visit)  + time_between_samples + yrs2initial + (1|Plot_ID), data = lf_long_join %>% filter(lifeform == "Shrubs") %>% filter(ecoregion == "Sierra Nevada"), family = tweedie)

plotResiduals(m_shrub) 
summary(m_shrub)

#model grams change over time
m_gram = glmmTMB(cover~ visit   + time_between_samples + yrs2initial + (1|Plot_ID), data = lf_long_join %>% filter(lifeform == "Grams") %>% filter(ecoregion == "Sierra Nevada"), family = tweedie)

plotResiduals(m_gram) 
summary(m_gram)

#model forbs change over time
m_forb = glmmTMB(cover ~ visit   + time_between_samples + yrs2initial + (1|Plot_ID), data = lf_long_join %>% filter(lifeform == "Forbs") %>% filter(ecoregion == "Sierra Nevada"), family = tweedie)

plotResiduals(m_forb) 
summary(m_forb)

```

Perform emmeans comparisons
```{r}
em_shrub_sim = emmeans(m_shrub_sim, specs = revpairwise ~ visit , data = lf_long_join %>% filter(lifeform == "Shrubs") %>% filter(ecoregion == "Klamath Mnts."), type = "response")

em_forb_sim = emmeans(m_forb_sim, specs = revpairwise ~ visit , data = lf_long_join %>% filter(lifeform == "Forbs") %>% filter(ecoregion == "Klamath Mnts."), type = "response")

em_gram_sim = emmeans(m_gram_sim, specs = revpairwise ~ visit , data = lf_long_join %>% filter(lifeform == "Grams") %>% filter(ecoregion == "Klamath Mnts."), type = "response")

em_shrub = emmeans(m_shrub, specs = revpairwise ~ visit , data = lf_long_join %>% filter(lifeform == "Shrubs") %>% filter(ecoregion == "Sierra Nevada"), type = "response")

em_forb = emmeans(m_forb, specs = revpairwise ~ visit , data = lf_long_join %>% filter(lifeform == "Forbs") %>% filter(ecoregion == "Sierra Nevada"), type = "response")

em_gram = emmeans(m_gram, specs = revpairwise ~ visit , data = lf_long_join %>% filter(lifeform == "Grams") %>% filter(ecoregion == "Sierra Nevada"), type = "response")


#combine emmeans into table
bind_rows(data.frame(em_shrub_sim$contrasts), data.frame(em_forb_sim$contrasts), 
          data.frame(em_gram_sim$contrasts), data.frame(em_shrub$contrasts), data.frame(em_forb$contrasts), data.frame(em_gram$contrasts))
```

